|

### 2023-07-17 paper/03-methods.Rmd
fixed abbreviations (#14)
updated fixed effects paragraph (#11)

*   The distance between pass rusher $j$ and the quarterback at frame $t$ during play $i$ is 
\begin{equation*}
s_{ij}(t) = \sqrt{(x_{ijt} - x^{QB}_{it})^2 + (y_{ijt} - y^{QB}_{it})^2}.
\end{equation*}
*   The velocity at which pass rusher $j$ is moving towards the quarterback at frame $t$ during play $i$ is 
\begin{equation*}
v_{ij}(t) = s'_{ij}(t) = \frac{ds_{ij}(t)}{dt}.
\end{equation*}
*   The STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation*}
\text{STRAIN}_{ij}(t) = \frac{v_{ij}(t)}{s_{ij}(t)}.
\end{equation*}

@@

As for the fixed-effects about pass rusher $j$ in play $i$, we include a variety of features that likely contribute to variation in STRAIN.
First, we adjust for the position of both the pass rusher and nearest blocker to account for any positional effects.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
These are encoded as indicator variables with defensive ends and tackles as the reference levels for the pass rushers and blockers, respectively.
We also adjust for the number of pass blockers on the play, since teams may decide to employ a more protective scheme that could lower the observed STRAIN.
Finally, we control for play-context covariates with respect to the offensive team.
These include the current down (first, second, third, fourth, or two-point conversion), yards to go for a first down, and current yardline (i.e., distance from the possession team's goal line).
[TODO: add explanation why play context is a proxy for time]

>>>>

*   The distance between pass rusher $j$ and the quarterback at frame $t$ during play $i$ is 
\begin{equation*}
s_{ij}(t) = \sqrt{(x_{ijt} - x^{QB}_{it})^2 + (y_{ijt} - y^{QB}_{it})^2}.
\end{equation*}
*   The velocity at which pass rusher $j$ is moving towards the quarterback at frame $t$ during play $i$ is 
\begin{equation*}
v_{ij}(t) = s'_{ij}(t) = \frac{ds_{ij}(t)}{dt}.
\end{equation*}
*   The STRAIN for pass rusher $j$ at frame $t$ during play $i$ is 
\begin{equation*}
\text{STRAIN}_{ij}(t) = \frac{- v_{ij}(t)}{s_{ij}(t)}.
\end{equation*}

@@

As for the fixed-effects about pass rusher $j$ in play $i$, we include a variety of features that likely contribute to variation in STRAIN.
First, we adjust for the position of both the pass rusher and nearest blocker to account for any positional effects.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
These are encoded as indicator variables with defensive ends and tackles as the reference levels for the pass rushers and blockers, respectively.
We also adjust for the number of pass blockers on the play, since teams may decide to employ a more protective scheme that could lower the observed STRAIN.
Finally, we control for play-context covariates with respect to the offensive team.
These include the current down (first, second, third, fourth, or two-point conversion), yards to go for a first down, and current yardline (i.e., distance from the possession team's goal line). We account for play-context since these variables impact a team's designed play, which may result in a play with low or high STRAIN regardless of the pass rusher's role. For instance, a team may call a short pass that is intended to be thrown early which could limit the potential STRAIN on a play. Or a team may need to throw a deep pass which would require more time and create more STRAIN potential. We do not account for time directly in the model due to the concern that the time it takes for a quarterback to throw the ball is itself a function of both the play call and pressure from pass rushers. Thus, since we do not know the designed play call, we condition on the play context to adjust for play-level differences attributing to a pass rusher's STRAIN.
